---
title: "covid 19"
author: "Taylor Stohl"
date: "Summer 2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata$sex <-as.factor(kRdata$sex)
kRdata$result <- as.factor(kRdata$result)
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

# Methods 

To change the category sex and result to factor variables, appropriate R chunks were added in. Afterwards, sex became a factor variable with two categories:female and male. Result, became a factor variable with two categories as well: alive and dead.


Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

`
```{r}
dd2 <- kRdata %>% group_by(result,sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
 
```{r}
barchartGC(~result + sex,data=kRdata)
barchartGC(~result + sex,data=kRdata, type="percent")

``` 

based on what can be seen in the percentage plot obove, it appears that approximately 60% of males died as a result of covid-19 related issues, whereas only about 40% of female participants died for the same reasons.



  
## Numerical results


```{r}
table1 <- xtabs(~result + sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages. The percentage of men that dies due to covid-19 related issues appears to be 64%, and the deaths among women for the same reason was only at 36%. 
Men died approximately twice as frequently as the women did for covid-19 related issues.



## Inferential results 




## Conclusions

